Algorithm explanation

### 2013-08-21 article/algo.tex
Adding example explanation

>>>>

\paragraph {Tree interpolation}
First, consider the following example.

\begin{align*}
x \leq 1 & \implies A(x) \\
y \leq 2 & \implies B(y) \\
A(x) \wedge B(y) & \implies 3x+2y \leq 10
\end{align*}

In this example, we would like to compute linear arithmetic
formulas for the predicate variables $A$ and $B$.
They appear only once on both left-hand and right-hand globally in the problem.

We try to obtain a simple solution that is expressed by linear expression
templates, whose coefficients and constants are parameterized. Assume
that they have following predicate templates.

\begin{align*}
A(x) : c_1 x + c_2 \leq 0 \\
B(y) : c_3 y + c_4 \leq 0
\end{align*}

By naturally encoding the first Horn clause with Farkas's lemma,
we obtain a linear constraint
$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will make a valid $A(x)$
for the first clause. In the same manner, we obtain
$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
from the second clause, and
$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
from the third.

One of the models satisfying all the constraints above is
$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
By assigning them to the template, we obtain the following solution.

\begin{align*}
A(x) : 3 x - 3 \leq 0 \\
B(y) : 2 y - 4 \leq 0
\end{align*}

\paragraph {DAG Horn graph solving}
Second, we explain a case that some predicate variables appear
multiple times on left-hand side in different Horn clauses.

\begin{align*}
x \leq 1 & \implies A(x) \\
A(x) \wedge A(y) & \implies x+y \leq 2
\end{align*}

Assigning a template
$A(x) : c_1 x + c_2 \leq 0$, we obtain the same
constraint for the first clause as the previous example.

\begin{align} \label{eq:constr1}
\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
\end{align}

Now we have $A$'s appearance twice on the second clause, and we try to
solve first by a weaker constraint, which we give allow different
assignments for multiple occurrence. Virtually it is equivalent to
solve the following problem with the templates.

\begin{align*}
\text{Problem:} \\
x \leq 1 & \implies A_1(x) \\
x \leq 1 & \implies A_2(x) \\
A_1(x) \wedge A_2(y) & \implies x+y \leq 2 \\
\\
\text{Templates:} \\
A_1(x) \text{ : } & c_{1,1} x + c_{2,1} \leq 0 \\
A_2(x) \text{ : } & c_{1,2} x + c_{2,2} \leq 0 \\
A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
\end{align*}

By this, duplicating the constraint ~\ref{eq:constr1}, we obtain the
whole constraint as follows.

\begin{align*}
& \lambda_{1,1} = c_{1,1} \wedge - \lambda_{1,1} \geq c_{2,1} \wedge \lambda_{1,1} \geq 0 \wedge \\
& \lambda_{1,2} = c_{1,2} \wedge - \lambda_{1,2} \geq c_{2,2} \wedge \lambda_{1,2} \geq 0 \wedge \\
& c_{1,1} = \lambda_2 \wedge c_{1,2} = \lambda_2 \wedge c_{2,1} + c_{2,2} \geq 2 \lambda_2 \wedge \lambda_2 > 0
\end{align*}

This constraint is satisfiable and one model is
\[ ( c_{1,1}, c_{1,2}, c_{2,1}, c_{2,2}, \lambda_{1,1}, \lambda_{1,2}, \lambda_2 ) =
( 1, -1, 1, -1, 1, 1, 1 ) \] and the solution then becomes
$A(x) : (x -1 \leq 0) \wedge (x -1 \leq 0)$.

Since the weak constraint is satisfiable, we now refine the constraint
to obtain the better solution. That is, simply eliminate the
duplication to build the following constraint for the original template.

\begin{align*}
& \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \\
& c_1 = \lambda_2 \wedge c_1 = \lambda_2 \wedge 2 c_2 \geq 2 \lambda_2 \wedge \lambda_2 > 0
\end{align*}

Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.

\paragraph {Disjunctive Horn graph solving}
Finally, we explain a case that an input set of Horn clauses contain
disjunctions on left-hand side.

\begin{align*}
x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
A(x) \wedge B(x) & \implies 1 \leq 0
\end{align*}
